\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{enumerate}
\usepackage[]{graphicx}
\usepackage{color,subfigure}
\usepackage{multicol}
\usepackage{float}
\usepackage[all]{xypic}
\usepackage[colorlinks=true,citecolor=cyan,linkcolor=magenta]{hyperref}
\usepackage{colonequals}

\usepackage{fancyhdr, lastpage}
\pagestyle{fancy}
\fancyfoot[C]{{\thepage} of \pageref{LastPage}}


\DeclareMathOperator{\mSpec}{mSpec}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Ass}{Ass}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\height}{height}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ann}{ann}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\M}{M}
\DeclareMathOperator{\Tor}{Tor}
\DeclareMathOperator{\Aut}{Aut}
%\DeclareMathOperator{\gcd}{gcd}

\def\ra{\rightarrow}
\newcommand{\m}{\mathfrak{m}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\ov}{\overline}

\title{}
\date{\vspace{-0.5in}}

\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother

\theoremstyle{definition}
\newtheorem{problem}{Problem}
\newtheorem*{solution}{Solution}


\title{Homework 1}

\begin{document}

\thispagestyle{fancy}
\pagestyle{fancy}
\rhead{UNL}
\lhead{Math 817}

\vspace{3em}

\begin{center}
	{\LARGE  Some old qualifying exam questions}
\end{center}


Here are some old qualifying exam problems you are already ready to solve.


\vspace{2em}




\begin{problem}
Let $H$ be a subgroup of a group $G$. Show that $[G,G] \leq H$ if and only if $H$ is a normal subgroup of $G$ and $G/H$ is abelian.
\end{problem}


\begin{problem}
Let $G$ be a group and $H$ a subgroup of $G$. The centralizer of $H$ in $G$ is the set of elements of $G$ that commute with each element of $H$:
  $$
  C_G(H) =: \{g \in G \mid gh = hg \text{ for all $h \in H$ }\}.
  $$
  Prove that if $H$ is normal in $G$, then $C_G(H)$ is a normal subgroup of $G$ and that $G/C_G(H)$ is isomorphic to a subgroup of the automorphism group of $H$.
\end{problem}



\begin{problem}
Let $G$ be a group (not necessarily finite) and $H$ a nonempty subset of $G$ that is closed under multiplication. Suppose that for all $g \in G$ we have $g^2 \in H$.
\begin{enumerate}[(a)]
	\item Show that $H$ must be a subgroup of $G$.
	\item Show that $H$ must be a normal subgroup of $G$.
	\item Show that $G/H$ must be abelian.
\end{enumerate}
\end{problem}


\begin{problem}
Fix a prime number $p$, and let $A$ denote the abelian group of all complex roots of unity whose orders are powers of $p$; that is,
$$
A= \{ z \in  \C \mid z^{p^n} =1 \text{ for some integer $n \geq 1$}\}. 
$$
Prove the following statements.
\begin{enumerate}[(a)]
\item Every non-trivial subgroup of $A$ contains the group of $p$-th roots of unity.
\item Every proper subgroup of $A$ is cyclic.
\item  If $B$ and $C$ are subgroups of $A$, then either $B \subseteq C$ or $C \subseteq B$.
\item For each $n \geq 0$, there exists a unique subgroup of $A$ with $p^n$ elements.
\end{enumerate}
\end{problem}

\begin{problem} Let $G$ be a group and $H$ and $K$ subgroups. Recall that $HK$ is the subset of $G$ defined as ${HK = \{hk \mid h \in H,k \in K\}}$.
\begin{enumerate}[(a)]
\item  Prove $HK=KH$ if and only if $HK$ is a subgroup of $G$.
\item Give an example (with justification) where $HK$ is not a subgroup.
\end{enumerate}
\end{problem}


\begin{problem} Let $G$ be a finite group.
\begin{enumerate}[(a)]
\item If $N$ is a normal subgroup of $G$ and $\#N = 2$, prove that $N$ is contained in the center $Z(G)$ of~$G$.
\item Suppose that $\# Z(G)$ is odd and that $G$ contains a non-trivial simple\footnote{$H$ is \textbf{simple} provided it has no normal subgroups
  other than $\{e\}$ and $H$ itself.} subgroup $H$ with ${[G : H] = 2}$. Prove that $H$ is the only non-trivial proper normal subgroup of $G$. 
\end{enumerate}
\end{problem}

\end{document}