\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm,enumerate}
\usepackage[]{graphicx}
\usepackage{color,subfigure}
\definecolor{scarlet}{rgb}{0.81,0,0}
\usepackage{multicol}
\usepackage{float}
\usepackage[all]{xypic}
\usepackage[colorlinks=true,citecolor=scarlet,linkcolor=scarlet]{hyperref}
\usepackage{colonequals}

\usepackage{fancyhdr, lastpage}
\pagestyle{fancy}
\fancyfoot[C]{{\thepage} of \pageref{LastPage}}



\DeclareMathOperator{\mSpec}{mSpec}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Ass}{Ass}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\height}{height}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ann}{ann}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\coker}{coker}
%\DeclareMathOperator{\ker}{ker}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\M}{M}
\DeclareMathOperator{\Tor}{Tor}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\ch}{char}
\DeclareMathOperator{\Aut}{Aut}
%\DeclareMathOperator{\dim}{dim}

\DeclareMathOperator{\lcm}{lcm}

\def\ra{\rightarrow}
\newcommand{\m}{\mathfrak{m}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\ov}[1]{\overline{#1}}

\def\ov#1{\overline{#1}}


\title{}
\date{\vspace{-0.5in}}

\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother

\theoremstyle{definition}
\newtheorem{problem}{Problem}


\begin{document}

\thispagestyle{fancy}
\pagestyle{fancy}
\rhead{UNL $\mid$ Fall 2025}
\lhead{Introduction to Modern Algebra I}

\vspace{3em}

\begin{center}
	{\LARGE Problem Set 1 \\}
	Due Wednesday, September 3
\end{center}

\

\noindent
{\bf Instructions:}
You are encouraged to work together on these problems, but each student should hand in their own final draft, written in a way that indicates their individual understanding of the solutions. Never submit something for grading that you do not completely understand. You cannot use any resources besides me, your classmates, and our course notes.


I will post the .tex code for these problems for you to use if you wish to type your homework. If you prefer not to type, please  {\em write neatly}. As a matter of good proof writing style, please use complete sentences and correct grammar. You may use any result stated or proven in class or in a homework problem, provided you reference it appropriately by either stating the result or stating its name (e.g. the definition of ring or Lagrange's Theorem). Please do not refer to theorems by their number in the course notes, as that can change.


\

%\begin{problem}
%Prove that every group of order $4$ is abelian. Your proof should only use the definition of a group. 
%\end{problem}

\begin{problem}
Let $G$ be a group and $x \in G$ any element. 
Recall that $|x|$ denotes the {\em order} of $x$, defined to be the least integer $n \geqslant 1$ such that $x^n = e$; if no such integer exists, we say $|x| = \infty$.
Also, let $|G|$ denote the cardinality of $G$; note that $|G|$ is an element of $\{1, 2, 3, \ldots \} \cup \{\infty\}$.

\begin{enumerate}[(a)]

\item Prove that if $|x| = n$, then $e, x, \dots, x^{n-1}$ are all distinct elements of $G$.

\item Prove that if $|x| = \infty$, then $x^i \neq x^j$ for all positive integers $i \neq j$. 

\item Conclude\footnote{We will soon prove Lagrange's Theorem, which states that $|x|$ divides $|G|$ when $G$ is finite. For now, use the definitions and not this result that we have not yet established.} that $|x| \leqslant |G|$ in all cases. 
\end{enumerate}
\end{problem} 


\begin{problem}
	A group $G$ is called {\it cyclic} if it is generated by a single element. 
	
	\begin{enumerate}[(a)]
		\item Prove\footnote{Your proof will be very short, as you can use the fact that $x^ix^j = x^{i+j}$ without proof.} that any cyclic group is abelian. 
	
		\item Prove that $(\mathbb Q, +)$ is not a cyclic group.
	\item Prove that $\operatorname{GL}_2(\mathbb Z/2)$ is not cyclic.
	\end{enumerate}
\end{problem}


\begin{problem} Let $n\geq 2$, and consider\footnote{Note: If you are unsure which formulas about permutations require proof, please ask.} the symmetric group $S_n$.
\begin{enumerate}[(a)]
\item Let $\tau\in S_n$ be a permutation, and $(i_1 \, i_2 \, \cdots \, i_k)$ be a $k$-cycle. Show that 
\[ \tau (i_1 \, i_2 \, \cdots \, i_k)  \tau^{-1} (\tau(i_1) \, \tau(i_2) \, \cdots \, \tau(i_k)).\]
\item Show that $S_n$ is generated by $(12)$ and the $n$-cycle $(12 \cdots n)$.
\item Show that, if $n\geq 3$, then $Z(S_n) = \{e\}$.
\end{enumerate}
\end{problem}



\begin{problem}
	\begin{enumerate}[(a)]
\item Suppose the cycle type of $\sigma \in S_n$ is $m_1, m_2, \ldots, m_k$. Recall this means that $\sigma$ a product of disjoint cycles of lengths $m_1, m_2, \ldots, m_k$. Prove that $|\sigma| = \lcm(m_1, \ldots, m_k)$. 
\item Given an example of two permutations $\sigma, \tau \in S_n$ such that $|\sigma\tau| > \lcm(|\sigma|,|\tau|)$.
\end{enumerate}
\end{problem}


\end{document}