
\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath,framed}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}
\def\e{\varepsilon}
\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}
\def\ds{\displaystyle}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Fake Problem Set\\
Due: Never}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 




\begin{enumerate}


\item Let $[a,b]$ be a closed interval and $f$ be a continuous function on $[a,b]$. Show that the range of $f$ (that is, $\{ f(x) \ | \ x\in [a,b]\}$) is a closed interval.

\item Let $f$ and $g$ be functions defined on $\R$ and $a$ a real number. Assume that $f$ is differentiable at $a$ and $f(a) = f'(a)=0$.
\begin{enumerate}
\item Use the product rule to show that if $g$ is differentiable at $a$, then  $(fg)'(a)=0$.
\item Show that if $g$ is continuous at $a$, then $(fg)'(a)=0$.
\item Show that if $g$ is \emph{not} continuous at $a$, then $fg$ may not be differentiable at $a$.
\end{enumerate}

\

Problems that will only make sense after Tuesday:

\

\item Let $f:\R\to\R$ be differentiable  on $\R$, $f(0)=0$, and $f'(x)<1$ for all $x\in \R$. Show that $f(x)<x$ for all $x>0$.

\

\item A fixed point of a function is a function $x$ is a number $r$ such that $f(r)=r$. Show that if $f:\R\to \R$ is differentiable on $\R$ and $f'(x)\neq 1$ for all $x\in \R$, then $f$ has at most one fixed point.

\end{enumerate}



\end{document}

























\end{document}