\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}
\def\e{\varepsilon}
\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}
\def\ds{\displaystyle}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#3 \\
Due: Thursday, September 15 by 7 pm, on Canvas}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}
\item Show that, for any positive number $\e>0$, there is a natural number $n$ such that $0 <  \frac1n <\e $.

\

\item Show that the supremum of the set $\ds S= \left\{ 2 - \frac3n \ | \ n\in \N\right\}$ is $2$.

\

\item Let $r$ be any real number. Consider the set
\[ S_r = \{ q\in \Q \ | \ q < r\}.\]
Prove that the supremum of $S_r$ is $r$.

\

\item Find all real numbers $x$ that satisfy the given inequality, and express your answers as intervals or unions of intervals.
\begin{enumerate}
\item $|2x+7| < 13$.
\item $|2x+7| \leq 13$.
\item $|2x+7| > 13$.
\end{enumerate}

\

\item \begin{enumerate}
\item Use the Triangle Inequality and the Reverse Triangle Inequality\footnote{Hint: Write $2x+3 = (2x - 3) + 6 = (2x-3) - (-6)$} to show that for all real numbers $x$,
\[ |2x-3| - 6 \leq |2x+3| \leq |2x-3| + 6.\]
\item Find (with justification) a positive real number $\delta$ such that:\\
If $|2x-3| < \delta$, then $|4x^2-9|< \frac{1}{100}$.
\end{enumerate}



\end{enumerate}

\end{document}

























\end{document}