\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}
\def\e{\varepsilon}
\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}
\def\ds{\displaystyle}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#4 \\
Due: Thursday, September 22 by 7 pm, on Canvas}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}
 \item Let $\{b_n\}_{n=1}^\infty$ be a sequence and $L$ be a real number. Suppose that $b_n = L$ for all $n >  1000000$ (but we know nothing about the first $1000000$ values of $b_n$). Show that $\{b_n\}_{n=1}^\infty$ converges to $L$.
 
 \



\item Let $\{a_n\}_{n=1}^\infty$ be a sequence, and suppose that $\{a_n\}_{n=1}^\infty$  converges to $8$.
\begin{enumerate}
\item Show that there is some real number $N$ such that $a_n\in (5,11)$ for all natural numbers $n>N$.
\item Show that there is some real number $N$ such that $a_n < 8.001 $ for all natural numbers $n>N$.
\item Explain why $\{ n\in \N \ | \ a_n < 7\}$ is finite (has finitely many elements).
\end{enumerate}

\

\item Prove that\footnote{By $\sqrt{n}$, we mean the positive number whose square is $n$. Such a number exists by a proof similar to the one that $\sqrt{2}$~exists.} the sequence $\ds \left\{ \frac{1}{\sqrt{n}} \right\}_{n=1}^\infty$ converges to $0$.

\

\item Let $r$ be a real number. Prove that there is no ``next'' rational number: i.e., that the set of rational numbers greater than $r$ has no minimum element.

\

\item Let $S$ be a set of real numbers, and $T=\{ x^2 \ | \ x\in S\}$.
\begin{enumerate}
\item Prove that if $T$ is bounded above, then $S$ is bounded above.
\item Prove or disprove the converse to part (a).
\end{enumerate}



\end{enumerate}

\end{document}

























\end{document}