\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}
\def\e{\varepsilon}
\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}
\def\ds{\displaystyle}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#5 \\
Due: Thursday, September 29 by 7 pm, on Canvas}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}

\item Consider the sequence $\ds \left\{ \frac{2n+1}{5n-2} \right\}_{n=1}^\infty$.
Use the definition of converges (but not Theorem~10.2) to show that this sequence converges to $\ds \frac{2}{5}$.

\

\item Prove that the sequence \[ \left\{ \frac{ 8 n^2 - 5n +3} {4 n^2 + 1} \right\}_{n=1}^\infty\]
converges. (This includes finding to what it converges.) You should use Theorem~10.2, but carefully justify every step using the Theorem.

\

%\item Let $\{a_n\}_{n=1}^\infty$ and $\{b_n\}_{n=1}^\infty$ be two sequences and $L$ be a real number. Suppose that there is some $M\in \R$ such that for all $n>M$, we have $a_n=b_n$. Prove that $\{a_n\}_{n=1}^\infty$ converges to~$L$ if and only if $\{b_n\}_{n=1}^\infty$ converges to~$L$.

\item \begin{enumerate}
\item Prove part (1) of Theorem~10.2 in the special case $c=2$.
\item Prove part (1) of Theorem~10.2 for a general value of $c$.
\end{enumerate}
 
 \

\item Prove that the sequence $\{\sqrt{n}\}_{n=1}^\infty$ diverges.

\

\item Assume that $\{a_n\}_{n=1}^\infty$ converges to zero, and that $a_n\geq 0$ for all natural numbers $n$. Prove that $\{ \sqrt{a_n}\}_{n=1}^\infty$ converges to zero also.






\end{enumerate}

\end{document}

























\end{document}