@@ -0,0 +1,120 @@
\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}
\def\e{\varepsilon}
\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}
\def\ds{\displaystyle}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#6 \\
Due: Thursday, October 6 by 7 pm, on Canvas}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}

\item Let $\{a_n\}_{n=1}^\infty$ and $\{b_n\}_{n=1}^\infty$ be sequences. Either prove\footnote{You may use Theorem~10.2 whenever convenient, but make sure you are using something the Theorem says, and nothing it doesn't!} or give a counterexample to each of the following:
\begin{enumerate}
\item If $\{a_n^2\}_{n=1}^\infty$ is divergent, then $\{a_n\}_{n=1}^\infty$ is divergent.
\item If $\{a_n\}_{n=1}^\infty$ and $\{b_n\}_{n=1}^\infty$ both diverge, then $\{a_n + b_n\}_{n=1}^\infty$ also diverges.
\item If $\{a_n\}_{n=1}^\infty$ converges and $\{b_n\}_{n=1}^\infty$ diverges, then $\{a_n + b_n\}_{n=1}^\infty$ diverges.
\item Suppose also that $b_n\neq 0$ for all $n\in \N$. If $\{b_n\}_{n=1}^\infty$ converges to $0$, then $\displaystyle \left\{\frac{a_n}{b_n}\right\}_{n=1}^\infty$ diverges.
\end{enumerate}

\

\item Let $\{a_n\}_{n=1}^\infty$ be any bounded sequence (not necessarily convergent) and let $\{b_n\}_{n=1}^\infty$ be a sequence that converges to $0$. Prove\footnote{Note that Theorem 10.2 does not apply. Try a different one!} that $\{a_n b_n\}_{n=1}^\infty$ converges to $0$.

\

\item Use the definition to prove that the sequence $\{ -\sqrt{n} \}_{n=1}^\infty$ diverges to $-\infty$.


\

\item Prove that if a sequence $\{ a_n\}_{n=1}^\infty$ diverges to $-\infty$ then it is not bounded below. 

\

\item Let $\{a_n\}_{n=1}^\infty$ be a sequence with $a_n>0$ for all $n\in \N$. Prove that $\{a_n\}_{n=1}^\infty$ converges to $0$ if and only if $\displaystyle \left\{\frac{1}{a_n}\right \}_{n=1}^\infty$ diverges to $+\infty$. 





\end{enumerate}

\end{document}

























\end{document}