\documentclass[12pt]{amsart}


\usepackage{times}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,multicol,graphicx,framed}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\inv}{^{-1}}
\newcommand{\dabs}[1]{\left| #1 \right|}
\newcommand{\e}{\varepsilon}
\newcommand{\ds}{\displaystyle}

\DeclareMathOperator{\res}{res}

%\usepackage{times}

%\addtolength{\textwidth}{100pt}
%\addtolength{\evensidemargin}{-45pt}
%\addtolength{\oddsidemargin}{-60pt}

\pagestyle{empty}
%\begin{document}\begin{itemize}

%\thispagestyle{empty}




\begin{document}
	
	\thispagestyle{empty}
	
	\section*{Subsequences \S 2.3}

	\



\begin{framed}
\noindent \textsc{Definition 20.2:} A \textbf{subsequence} of a given sequence
  $\{a_n\}_{n=1}^\infty$ is any sequence of the form $\{a_{n_k}\}_{k=1}^\infty$
where 
$\{n_k\}_{k=1}^\infty$
is any strictly increasing sequence of natural numbers.\end{framed}

\

\begin{enumerate}
 \item \textbf{True or false; justify.}
\begin{enumerate} 
\item The sequence $\displaystyle \left\{ \frac{1}{3n+7} \right\}_{n=1}^\infty$ is a subsequence of the sequence $\displaystyle \left\{ \frac{1}{n} \right\}_{n=1}^\infty$.

\

\item The constant sequence $\displaystyle \left\{ \frac{1}{2} \right\}_{n=1}^\infty$ is a subsequence of the sequence $\displaystyle \left\{ \frac{1}{n} \right\}_{n=1}^\infty$.

\

\item The constant sequences $\{ -1 \}_{n=1}^\infty$ and  $\{ 1 \}_{n=1}^\infty$ are both subsequences of the sequence $\{ (-1)^n \}_{n=1}^\infty$.

\

\item The constant sequences $\{ -1 \}_{n=1}^\infty$ and  $\{ 1 \}_{n=1}^\infty$ are the only two subsequences of the sequence $\{ (-1)^n \}_{n=1}^\infty$.

\

\item The sequence $\{ \sin(\pi n) \}_{n=1}^\infty$ is a subsequence of $\{ \sin(n) \}_{n=1}^\infty$.

\end{enumerate}

\

\item Explain how the following Corollary follows from Theorem~20.4:

\

\noindent \textsc{Corollary 20.6:} Let $\{a_n\}_{n=1}^\infty$ be any sequence.
\begin{enumerate}
\item If there is a subsequence of $\{a_n\}_{n=1}^\infty$ that diverges, then the sequence $\{a_n\}_{n=1}^\infty$ diverges.
\item If there are two subsequences of $\{a_n\}_{n=1}^\infty$ that converge to different values, then $\{a_n\}_{n=1}^\infty$ diverges.
\end{enumerate}

\



\item Use Corollary 20.6 to give a quick proof that the sequence $\{ (-1)^n \}_{n=1}^\infty$ diverges.

\

\item \textbf{Prove or disprove:}
\begin{enumerate}
\item Every subsequence of a bounded sequence is bounded.

\

\item Every subsequence of a divergent sequence is divergent.

\

\item Every subsequence of a sequence that diverges to $-\infty$ also diverges to $-\infty$.




\end{enumerate}

\end{enumerate}






\end{document}
