\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}

\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 445 --- Problem Set \#2 \\
Due: Friday, September 8 by 7 pm, on Canvas}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}

\item Let $a,b,c$ be integers. Show that if $a$ and $b$ are coprime, $a$ divides $c$, and $b$ divides $c$, then $ab$ divides $c$.


\

\item Find all solutions to the equation $x^2 + [4]x = [5]$ in $\Z_8$ by trial and error (plugging in all possible values). Use this to find all integer solutions to $x^2+ 4x \equiv 5 \pmod 8$.

\

\item  Given integers $a_1,\dots,a_m$, the \textbf{greatest common divisor} of $a_1,\dots,a_m$ is the largest integer that divides all of them. 
\begin{enumerate}
%\item Show that $\gcd(a,b,c) = \gcd(a,\gcd(b,c))$.
\item Compute $\gcd(12,18,42)$.
\item Prove or disprove: If $\gcd(a,b,c) = 1$, then some pair of the numbers $a,b,c$ is coprime.
\end{enumerate}

\

\item \emph{Use the methods we have developed in class} to solve the following:
\begin{enumerate}
\item Find all integer pairs $(x,y)$ such that $275x-126y=9$.
\item Find the inverse of $[126]$ in $\Z_{275}$.
\item Find the smallest positive integer $x$ such that 
\[ x\equiv 7 \pmod{126} \quad\text{and}\quad x\equiv 8 \pmod{275}.\]
\end{enumerate}

\

\item  Solving linear equations in $\Z_n$: Let $a,b,n$ be integers, with $n>0$.
\begin{enumerate}
\item Show that $[a] x = [b]$ has a solution $x$ in $\Z_n$ if and only if $\gcd(a,n)$ divides $b$.
\item Show that if $[a] x = [b]$ has a solution $x$ in $\Z_n$, then there are exactly $\gcd(a,n)$ distinct solutions.
\item Solve the equation $[20] [x] + [17] = [29]$ in $\Z_{36}$.
\end{enumerate}


\end{enumerate}

\noindent  \hrulefill

\noindent The remaining problems are only required for Math 845 students, though all are encouraged to think about them.

\

\begin{enumerate}\setcounter{enumi}{5}



\item Solve the equation $8x + 25y + 15z = 19$ over $\Z$.

%\item

\end{enumerate}

\end{document}

























\end{document}