 \documentclass{amsart}
\usepackage{pdfpages}
% \pagestyle{empty}
\usepackage{amscd}
\usepackage{amssymb}
%\usepackage[all, knot]{xy}



\usepackage[top=1in, bottom=1in, left=1.1in, right=1.1in]{geometry}

%\setlength{\oddsidemargin}{-.10in}
%\setlength{\evensidemargin}{-.10in}
%\setlength{\textwidth}{5.5in}
%\setlength{\topmargin}{-.250in}
%\setlength{\headheight}{0in}
%\setlength{\headsep}{0in}
%\setlength{\topskip}{0in}
%\setlength{\textheight}{9.5in}
%\parindent = 0in
\font\bigbf = cmbx10 scaled \magstep1
\font\medbf = cmbx10 scaled \magstephalf
\begin{document}
 
%\magnification \magstep1
%\parindent = 0pt
%\nopagenumbers
%\voffset = -.5truein
%\vsize = 10 truein
%\baselineskip = 1.5 \baselineskip
%\font\bigbf = cmbx10 scaled \magstep1
%\font\medbf = cmbx10 scaled \magstephalf








\centerline{\bigbf Algebra II: Math 902 -- Section 001}
\centerline{\bigbf Spring Semester 2022}
\centerline{\bigbf MWF 11:30pm -- 12:20pm }
\centerline{\bigbf Burnett 232}



\bigskip

\noindent
{\bf Instructor:}  Jack Jeffries

\noindent
{\bf Office:} 325 Avery Hall

\noindent
{\bf email:} jack.jeffries@unl.edu

\

\noindent
{\bf Office Hours:} By appointment, or just drop in. You are most likely to find me on campus on Mondays, Wednesdays, and Fridays, but I can also meet on zoom on other days.


\

\noindent
{\bf Text:} None required. Notes will be online.


\

\noindent
{\bf Course Description:} 
In this class we will cover the basics of commutative algebra. Topics will include finiteness conditions, algebraic varieties and the Nullstellensatz, local rings, associated primes and primary decomposition, Krull dimension, Noether normalization, Krull's height theorem, Hilbert functions, regular and normal rings.

\


\noindent {\bf Assessments:} We will have 5 or 6 written homework assignments, for which you are strongly encouraged to work in groups. We will also two take-home exams, for which you are expected to work alone. 

\


{\hfill \begin{tabular}{|l|l|}
\hline
Component & Value \\
\hline
Problem Sets & 60\% \\
\hline
Midterm Exam & 20\% \\
\hline
Final Exam & 20\% \\
\hline
\end{tabular} \hfill}

\

\

\noindent {\bf Students With Disabilities:} Students with disabilities are encouraged to contact the instructor for a confidential discussion of their
individual needs for academic accommodation. It is the policy of the University of Nebraska-Lincoln to provide flexible and individualized accommodation to
students with documented disabilities that may affect their ability to fully participate in course activities or to meet course requirements. To receive
accommodation services, students must be registered with the Services for Students with Disabilities office, 132 Canfield Administration, 472-3787 voice or TTY,
{\tt http://www.unl.edu/ssd.}


\


\noindent {\bf Department Grading Policy:} Students who believe their academic evaluation has been prejudiced or capricious have recourse for appeals to (in
order) the instructor, the department vice chair, the department chair, the departmental appeals committee, and the college appeals committee. 

\

\noindent
{\bf Academic Honesty:}
Academic honesty is essential to the existence and integrity of an academic institution. The responsibility for maintaining that integrity is shared by all members of the academic community. The University's Student Code of Conduct addresses academic dishonesty. Students who commit acts of academic dishonesty are subject to disciplinary action and are granted due process and the right to appeal any decision.


\

\noindent
{\bf Face coverings:} Students are required to comply with all university face covering policies throughout the semester, and strongly recommended to comply with all local health guidelines. If anyone has a documented need for face coverings to be required in their classes, for example a member of your household is unable to be vaccinated or has a health condition that makes vaccines less effective for them, please contact me ASAP.


\vfill
\pagebreak
\smallskip
%\includepdf{UNLFace.pdf}



\end{document}


