\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=.7in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}

\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#10\\
Due: Wednesday, December 1 by 5 pm}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

Please write neatly, using complete sentences and correct
puctuation. Label the problems clearly. 


\



\begin{enumerate}


\item Show that\footnote{Hint: Break into three cases.}
\[ f(x) = \begin{cases} x^2-2x& \text{if} \ x\geq 2 \\ 2x-4 &\text{if} \ x<2 \end{cases}\]
is differentiable at all values of $x$, and compute the derivative.

\

\

\item Show that the function 
\[ f(x) = \begin{cases} x^2  & \text{if} \ x\in \Q \\ 0&\text{if} \ x\notin \Q \end{cases}\]
is differentiable at $x=0$, and not differentiable at any other value of $x$.

\

\

\item Let $f$ and $g$ be functions and $a$ be a real number. Assume that $f(a) = f'(a) = 0$ and that $g$ is continuous at $a$. 
\begin{enumerate}
\item Prove\footnote{Note: We are not assuming that $g$ is differentiable, so we cannot apply the product rule.} that $(fg)'(a) = 0$.
\item Is the statement still true if we do not assume that $g$ is continuous at $a$? Prove or disprove it.

\end{enumerate}
\end{enumerate}

\end{document}

























\end{document}