\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=.7in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}

\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#11\\
Due: Wednesday, December 8 by 5 pm}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

Please write neatly, using complete sentences and correct
puctuation. Label the problems clearly. 


\



\begin{enumerate}




\item Let $p(x)$ be a polynomial function.
\begin{enumerate}
\item Use Rolle's Theorem to show that if $p(x)$ has $n$ real roots, then $p'(x)$ has at least $n-1$ real roots.
\item Use Rolle's Theorem and induction to show that if $p(x)$ is not a constant function, then the number of real roots of $p(x)$ is at most the degree of $p$.
\end{enumerate}

\

\item Let $f:\R \to \R$ be differentiable on $\R$, and assume that $f(0)=0$, and that $f'(x)<1$ for all $x\in \R$.
\begin{enumerate}
\item Show that $f(x)<x$ for all $x > 0$.
\item Does the same equality necessarily hold for $x<0$? Justify your answer.
\end{enumerate}

\

\item A \emph{fixed point} of a function $f$ is a number $x$ such that $f(x)=x$. Prove that if $f$ is differentiable on $\R$, and $f'(x)\neq 1$ for all $x\in \R$, then $f$ has at most one fixed point.
\thispagestyle{empty}
\end{enumerate}

\end{document}

























\end{document}