\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}

\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#4\\
Due: Wednesday, September 22 by 5 pm}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}




\item 
Given any two real numbers $x$ and $y$, $\max \{x, y\}$ refers to the larger of the two numbers $x$ and
$y$; that is, $\max \{x,y\}$ is $x$ if $x \geq y$ and otherwise it is~$y$.
Similarly, $\min \{x, y\}$ refers to the smaller of the two numbers $x$ and
$y$; that is, $\min\{x,y\}$ is $x$ if $x \leq y$ and otherwise it is~$y$.


\begin{enumerate}
	
	\item
	Prove that for all real numbers $x$ and $y$
	$$
	\max \{x,y\} = \frac{x+y + |x-y|}{2}.
	$$
	
	\item Find a similar formula for $\min \{x,y\}$ and prove that your
	formula is correct.
	
\end{enumerate}

\

 \item Prove, using the formal definition of convergence, that the sequence $\{\frac{1}{\sqrt{n}}\}_{n=1}^\infty$ converges to~$0$. 
(At the risk of being overly pedantic, for $n \in \N$, by $\sqrt{n}$
we mean the unique positive real number whose square is $n$. Such a number exists by the Completeness Axiom, which proved in detail in the case $n = 2$.)

\

\item Let $\{a_n\}_{n=1}^\infty$ and $\{b_n\}_{n=1}^\infty$ be two sequences and $L$ be a real number. Suppose that there is some $M\in \R$ such that for all $n>M$, we have $a_n=b_n$. Prove that $\{a_n\}_{n=1}^\infty$ converges to~$L$ if and only if $\{b_n\}_{n=1}^\infty$ converges to~$L$.

\

\item Do problem \#5 on page 91 of the textbook.

\

\item Do problem \#6 on page 91 of the textbook.

\

\item Let $\{a_n\}_{n=1}^\infty$ be a sequence, and $K,L$ be real numbers. Suppose that for all $n\in \N$, $a_n\geq K$, and that $\{a_n\}_{n=1}^\infty$ converges to $L$. Prove that $L\geq K$. 

\



\end{enumerate}

\end{document}

























\end{document}