\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}

\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#5 \\
Due: Wednesday, September 29 by 5pm}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}

\item Prove that the sequence \[ \left\{ \frac{ 8 n^2 - 5n +3} {4 n^2 + 1} \right\}_{n=1}^\infty\]
converges. (This includes finding to what it converges.) You should use Theorem 12.2, but carefully justify every step using the Theorem.

\

\item Prove that the sequence $\{\sqrt{n}\}_{n=1}^\infty$ diverges.

\

\item Assume that $\{a_n\}_{n=1}^\infty$ converges to zero, and that $a_n\geq 0$ for all natural numbers $n$. Prove that $\{ \sqrt{a_n}\}_{n=1}^\infty$ converges to zero also.

\

\item Assume that $\{a_n\}_{n=1}^\infty$ converges to $L$, and that $a_n\geq 0$ for all natural numbers $n$. Prove\footnote{Use an old HW problem to explain why $L\geq 0$. Use the previous problem to deal with the case $L=0$. For $L>0$, use $(\sqrt{a_n}-\sqrt{L})(\sqrt{a_n}+\sqrt{L})=a_n-L$ to deduce that \[\sqrt{a_n}-\sqrt{L} \leq \frac{ |a_n - L|}{\sqrt{L}}.\]} that $\{ \sqrt{a_n}\}_{n=1}^\infty$ converges to $\sqrt{L}$.

\


\item Let $\{a_n\}_{n=1}^\infty$ converge to zero, and let $\{b_n\}_{n=1}^\infty$ be any bounded sequence (that may or may not be convergent). Prove that $\{a_n b_n\}_{n=1}^\infty$ converges to zero.

\

\item Let $\{a_n\}_{n=1}^\infty$ be a sequence such that $a_n>0$ for all $n$. Show that $\{a_n\}_{n=1}^\infty$ diverges to $\infty$ if and only if $\{\frac{1}{a_n}\}_{n=1}^\infty$ converges to $0$.





\end{enumerate}

\end{document}

























\end{document}