\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=.7in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}

\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#9\\
Due: Wednesday, November 10 by 5 pm}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}


\item Prove that the function $f(x) = \sqrt{|x|}$ is continuous at every real number $a$ using just the $\e-\d$ definition of continuity. 
  
  \
  
\item Let $f(x)$ be the function with domain all of $\R$ defined by
  $$
  f(x) = \begin{cases}
    2 + x, & \text{if $x \leq 2$ and} \\
    2x - 5, & \text{if $x > 2$.} \\
  \end{cases}
  $$
  Prove $f$ is not continuous at $x = 2$ using just the $\e-\d$  definition of continuity.

\
 
\item Prove $\sqrt{\sqrt{x^2 +1} + x^4 + 1}$ is continuous on all of  $\R$. You may use any theorems we've covered in class, but be sure to use them carefully.

\


\item Suppose $f$ is a function that is continuous on all of $\R$. Prove\footnote{Hint: You may want to use that
  every real number is the limit of some sequence of rational numbers.} that if $f(q) = 0$ for all $q \in \Q$, then $f(x) = 0$ for all $x$. 


\end{enumerate}

\end{document}

























\end{document}