\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm,enumerate,bbm}
\usepackage[]{graphicx}
\usepackage{color,subfigure}
\definecolor{scarlet}{rgb}{0.81,0,0}
\usepackage{multicol}
\usepackage{float}
\usepackage[all]{xypic}
\usepackage[colorlinks=true,citecolor=scarlet,linkcolor=scarlet]{hyperref}
\usepackage{colonequals}

\usepackage{fancyhdr, lastpage}
\pagestyle{fancy}
\fancyfoot[C]{{\thepage} of \pageref{LastPage}}



\DeclareMathOperator{\mSpec}{mSpec}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Ass}{Ass}
\DeclareMathOperator{\Supp}{Supp}
\DeclareMathOperator{\height}{height}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ann}{ann}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\coker}{coker}
%\DeclareMathOperator{\ker}{ker}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\M}{M}
\DeclareMathOperator{\Tor}{Tor}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\ch}{char}
\DeclareMathOperator{\Aut}{Aut}

%\DeclareMathOperator{\dim}{dim}

\DeclareMathOperator{\lcm}{lcm}

\def\ra{\rightarrow}
\newcommand{\m}{\mathfrak{m}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\ov}[1]{\overline{#1}}
\newcommand{\norm}{\trianglelefteq}

\def\ov#1{\overline{#1}}


\title{}
\date{\vspace{-0.5in}}

\makeatletter
\g@addto@macro\@floatboxreset\centering
\makeatother

\theoremstyle{definition}
\newtheorem{problem}{Problem}


\begin{document}

\thispagestyle{fancy}
\pagestyle{fancy}
\rhead{UNL $\mid$ Spring 2026}
\lhead{Introduction to Modern Algebra II}

\vspace{3em}

\begin{center}
	{\LARGE Problem Set 13 \\}
	Due never\end{center}

\

\noindent
{\bf Instructions:}
You are encouraged to work together on these problems, but each student should hand in their own final draft, written in a way that indicates their individual understanding of the solutions. Never submit something for grading that you do not completely understand. You cannot use any resources besides me, your classmates, and our course notes.


I will post the .tex code for these problems for you to use if you wish to type your homework. If you prefer not to type, please  {\em write neatly}. As a matter of good proof writing style, please use complete sentences and correct grammar. You may use any result stated or proven in class or in a homework problem, provided you reference it appropriately by either stating the result or stating its name (e.g. the definition of ring or Lagrange's Theorem). Please do not refer to theorems by their number in the course notes, as that can change.



\begin{problem}
Show that $\Q(\sqrt{2+\sqrt{2}})/\Q$ is a Galois extension of degree 4 with Galois group that is a cyclic group of order $4$.	
\end{problem}

\


\begin{problem}
	Let $F \subseteq L$ be Galois field extension of degree 45. Prove there exists a unique intermediate field $E$ such that $[E : F] = 5$.
\end{problem}

\

\begin{problem}
Let $F$ be a field with $p^6$ elements for some prime $p$. Determine the number of subfields of $F$, and compute the number of elements in each such subfield.
\end{problem} 

\

\begin{problem} Let $F$ be a field of characteristic $p > 0$, $a \in F$, and consider the polynomial $f(x)=x^p -x-a \in F[x]$.

  \begin{enumerate}[(a)]

    \item Prove\footnote{Hint: If $\alpha$ is a root of
      $f(x)$,  consider $\alpha + j$ for $j \in \Z/p$.} that $f(x)$ is either irreducible over $F$ or it splits
      into distinct linear factors over $F$.

      \item Suppose $f(x)$ is irreducible over $F$ and let $L$ be a
        splitting field of $f$ over $F$.  Prove that the Galois group of $L$ over $F$ is cyclic.

\end{enumerate}
\end{problem} 

\


\begin{problem}
  Let $p$ be any positive prime integer.
  \begin{enumerate}[(a)]
  \item Prove that if $p = k^2 + 1$ for some integer $k$, then $p$ is not an irreducible element of $\mathbb{Z}[i]$.
\item Prove $x^4 - p$ is irreducible\footnote{Hint: Find the degree of the splitting field of $x^4 - p$ over $\mathbb{Q}(i)$. }  in $\mathbb{Q}(i)[x]$.
\end{enumerate}
\end{problem} 

\

\begin{problem}  Let $F \subseteq L$ be a finite Galois field extension of degree $45$. Prove there exists a unique intermediate field
$E$ (i.e., $F \subseteq E \subseteq  L$) such that $[E : F] = 5$.
\end{problem} 




\end{document}