@@ -0,0 +1,120 @@
\documentclass{amsart}
\usepackage{amstext,amsfonts,amssymb,amscd,amsbsy,amsmath}
\usepackage{geometry}[margin=1in]
\pagestyle{empty}
%\parindent = 0in


\def\sol#1{{\bf Solution: } #1}
%\def\sol#1{}
\def\bl{\vskip .1in}
\def\star{${}^*$}
\def\cS{\mathcal S}
\def\cT{\mathcal T}
\def\cB{\mathcal B}
\def\e{\varepsilon}
\def\R{\mathbb R}
\def\N{\mathbb N}
\def\Q{\mathbb Q}
\def\Z{\mathbb Z}
\def\ds{\displaystyle}

\def\cC{\mathcal C}

\def\e{\epsilon}
\def\d{\delta}

\begin{document}



\begin{center}
{\large\bfseries
Math 325-002 --- Problem Set \#7 \\
Due: Thursday, October 27 by 7 pm, on Canvas}
\end{center}





{\bf Instructions:} You are encouraged to work together on these
problems, but each student should hand in their own final draft,
written in a way that indicates their individual understanding of
the solutions. Never submit something for grading
that you do not completely understand. 

If you do work with others, I ask that you write something along the
top like ``I collaborated with Steven Smale on problems 1 and 3''.
If you use a reference, indicate so clearly in your solutions. 
In short, be intellectually
honest at all times.

Please write neatly, using complete sentences and correct
punctuation. Label the problems clearly. 






\begin{enumerate}

\item For each of the following, give an explicit example as indicated; no proofs are necessary.
\begin{enumerate}
\item A sequence that has a subsequence that converges to $1$, another subsequence that converges to $2$, and a third subsequence that converges to $3$. 
\item A sequence that has one sequence that is monotone and converges to $0$ and another subsequence that is monotone and diverges to $+\infty$.
\end{enumerate}

\


\item Prove that if $\{a_n\}_{n=1}^\infty$ diverges to $+\infty$, then every subsequence of $\{a_n\}_{n=1}^\infty$ diverges to $+\infty$.

\

\item Define a sequence $\{a_n\}_{n=1}^\infty$ recursively by $a_1=2$ and $\displaystyle a_{n+1} = \frac{a_{n}}{2} + \frac{1}{a_{n}}$.
\begin{enumerate}
\item Prove that $a_n>0$ for all $n\in \N$.
\item Prove\footnote{Write $a_{n}-2$ in terms of $a_{n-1}$ and factor the expression.} that $a_n^2 \geq 2$ for all $n\in \N$.
\item Prove\footnote{Consider $a_{n+1}-a_n$ and use (b).} that the sequence is decreasing.
\item Show that the sequence is convergent.
\item Determine\footnote{If the sequence converges to $L$, explain why $L = \frac{L}{2} + \frac{1}{L}$, and solve for $L$.} what value the sequence converges to.
\end{enumerate}

\

\item Let $\{a_n\}_{n=1}^\infty$ be a sequence and $L$ a real number. Show that $\{a_n\}_{n=1}^\infty$ converges to $L$ if and only if both of the subsequences $\{a_{2k}\}_{k=1}^\infty$ and $\{a_{2k+1}\}_{k=1}^\infty$ converge to $L$.


\end{enumerate}

\end{document}

























\end{document}