\documentclass[12pt]{amsart}


\usepackage{times}
\usepackage[margin=.7in]{geometry}
\usepackage{amsmath,amssymb,multicol,graphicx,framed,enumitem}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\e}{\varepsilon}
\def\d{\delta}
\newcommand{\cts}{continuous\,}

\newcommand{\dabs}[1]{\left| #1 \right|}
\newcommand{\ds}{\displaystyle}
\usepackage[all, knot]{xy}
\usepackage{color,xcolor}
%\usepackage{times}

%\addtolength{\textwidth}{100pt}
%\addtolength{\evensidemargin}{-45pt}
%\addtolength{\oddsidemargin}{-60pt}

\pagestyle{empty}
%\begin{document}\begin{itemize}

%\thispagestyle{empty}




\begin{document}
	
	\thispagestyle{empty}
	
	\section*{Basics of Derivatives}
	
\begin{framed} 
\noindent \textbf{Definition:} Let $f$ be a function and $r$ be a real number. We say that $f$ is \emph{differentiable at $r$} if $f$ is defined at $r$ and the limit
\[ \lim_{x\to r} \frac{ f(x) - f(r) }{x-r}\]
exists. In this case, we call the limit \emph{the derivative of $f$ at $r$} and write $f'(r)$ for this limit. 
 \end{framed}
 
 
 \begin{enumerate}
 \item Use the definition to show that the derivative of a constant function is zero.
 
 \
 
 \item Use the definition to show that the function $f(x) = 2x-5$ is differentiable at $x=3$ and compute its derivative. 
 
 \
 
 \item Use the definition to show that the function $f(x)=|x|$ is \emph{not} differentiable at $x=0$.
 
 \
 
 \item Prove\footnote{Hint: Write down the function in the definition of derivative, and multiply both sides by $(x-r)$ and consider the limit.} that if $f$ is differentiable at $x=r$, then $f$ is continuous at $x=r$. 
 
 \
 
 \item Prove or disprove the converse of the previous statement.
 \end{enumerate}
 
 \
 
 
 \begin{framed} 
 \noindent \textbf{Theorem (Derivatives and algebra:} Let $f,g$ be functions that are differentiable at $x=r$, and $c$ be a real number. Then,
 \begin{enumerate}
 \item $f+g$ is differentiable at $x=r$ and $(f+g)'(r) = f'(r) + g'(r)$;
 \item $cf$ is differentiable at $x=r$ and $(cf)'(r) = c f'(r)$;
 \item $fg$ is differentiable at $x=r$ and $(fg)'(r) = f'(r) g(r) + f(r) g'(r)$.
   \end{enumerate}
 \end{framed}
 
 \
 
 \begin{enumerate}\setcounter{enumi}{5}
 \item Prove\footnote{You many want to use part (3) of the Theorem above.} that if $f(x)=x^n$, then $f$ is differentiable at any value of $x$ and $f'(x) = n x^{n-1}$ for every $n\in \N$.
 
 \
 
 \item Use the Theorem plus the previous problem to compute the derivative of $f(x) = 5x^7 - \sqrt{19} x^4$.
 
 \
 
 \item Prove the Theorem.
 
 \
 
 \end{enumerate}
 
 \newpage
 
	\section*{Derivatives and optimization}


 
 \begin{framed} 
 \noindent \textbf{Theorem:} Let $f$ be a function that is differentiable at $x=r$.
 \begin{enumerate}
 \item 
 If $f'(r) > 0$, then there is some $\delta>0$ such that 
 \begin{itemize}
 \item if $x\in (r,r+\delta)$ then $f(r) < f(x)$;
  \item if $x\in (r-\delta,r)$ then $f(x) < f(r)$.
  \end{itemize}
  
  \item  If $f'(r) < 0$, then there is some $\delta>0$ such that 
  \begin{itemize}
 \item if $x\in (r,r+\delta)$ then $f(r) > f(x)$;
 \item if $x\in (r-\delta,r)$ then $f(x) > f(r)$.
  \end{itemize}
  \end{enumerate}
  
  \
  
  \noindent \textbf{Corollary (Derivatives and optimization):}
  Let $f$ be a function that is continuous on a closed interval $[a,b]$. If $f$ attains a maximum or minimum value on $[a,b]$ at $r\in (a,b)$, and $f$ is differentiable at $r$, then $f'(r)=0$.
  
 \end{framed}
 
 \begin{enumerate}
 
 \item Find the values of $x$ on $[0,2]$ at which $f$ achieves its minimum and maximum values.
 
 \ 
 
 \item Explain why the Corollary follows from the Theorem.
 
 \
 
 
 \item Prove\footnote{Hint: Apply the definition of limit within the definition of derivative with a useful choice of $\e$.} part (1) of the Theorem.


\

 \item Prove part (2) of the Theorem.
 
\end{enumerate}

\
 
 \noindent A function $f$ is \emph{increasing} on an interval $(a,b)$ if for any $r,s\in (a,b)$ with $r<s$, we have $f(r)<f(s)$.
 
 \
  \begin{enumerate}\setcounter{enumi}{4}
 \item Prove or disprove: If $f'(r)>0$, then there is some $\delta>0$ such that $f$ is increasing on $(r-\delta,r+\delta)$.
\end{enumerate}

\end{document}
